### 2017-10-02 1/exams/m1.tex

\begin{center}
  {\Large {\bf CMSC 131A, Midterm 1 (Practice)}\\
    \ \\
    Fall 2017}
\end{center}

\begin{center}
\begin{tabular}{| c | c |}
  \hline
  Question & Points \\ \hline \hline
  1 & 10 \\ \hline
  2 & 10 \\ \hline
  3 & 10 \\ \hline
  4 & 20 \\ \hline
  5 & 10 \\ \hline
  6 & 20 \\ \hline \hline
  7 & 20 \\ \hline \hline
  Total: & 100\\
  \hline
\end{tabular}
\end{center}

@@

@@

\noindent
{\bf Problem 1 (10 points).}
%
For each of the following programs, write out each step of
computation. At each step, underline the expression being
simplified. Label each step as being ``arithmetic'' (meaning any
built-in operation), ``conditional'', ``plug'' (for plugging in an
argument for a function parameter), or ``constant'' for replacing a
constant with its value.

\noindent
{\bf Problem 1(a).}

\begin{verbatim}
(define R 9)
(define (w z) (* R z))
(add1 (cond [(= (w 2) 12) 4]
            [else 9]))
\end{verbatim}

@@

\noindent
{\bf Problem 1(b).}

\begin{verbatim}
(posn-x (cond [(= (string-length "hi") 2) (make-posn 3 4)]
              [else (make-posn 8 0)]))
\end{verbatim}

@@

\noindent
{\bf Problem 2 (10 points).}
%
The graphical text editor you developed in a recent assignment has
garned major attention by Bay-area venture capitalists who want to
base a start-up company around your editor.  They're ready to invest
to help you develop a full-featured prototype.  But before they write
a check, they want to make sure you can handle fancy operations for
expert text editor users.  They ask if you can implement a
\emph{transpose} operation, which will transpose the character to the
left of a cursor with the character to the right of the cursor.  To
satisfy your potential investors, you define the following function
for transposing characters at a given position in a string that has at
least 2 characters in it:
\begin{verbatim}
;; transpose : String Index -> String
;; Transpose characters at index i and i+1.
;; Assumes string has length >= 2 and 0<=i, i+1<=length.
(check-expect (transpose "ab" 0) "ba")
(check-expect (transpose "peice" 1) "piece")
(check-expect (transpose "student's" 7) "students'")
(define (transpose s i) ...)
\end{verbatim}

\noindent
Give a correct definition for \verb|transpose|.  (You only need to
provide code, not the design steps.)

\newpage 
\noindent
{\bf Problem 3 (10 points).}
%
A point in three dimensional space can be represented by three numbers
specifying the points distance from the $x$-axis, $y$-axis, and
$z$-axis.  The distance of a point $(x,y,z)$ can be computed as:
\[\sqrt(x^2+y^2+z^2)\]
For example, the distance of $(2,6,9)$ to the origin is $11$; the
distance of $(3,2,6)$ is $7$.  Assume the following data definition
and design a program for computing the distance of a given 3D
point.

\begin{verbatim}
;; A 3D is a (make-3d Number Number Number)
;; Interp: a point in three dimension space
(define-struct 3d (x y z))
\end{verbatim}

@@

\newpage
\noindent
{\bf Problem 4 (20 points).}  Growing tired of Space Invaders, you
decide to build a new video game.  Part of the game consists of flying
billiard balls that move in straight lines at varying velocities until
hitting other balls or bouncing off walls and other obstacles.  After
making some sketches, you decide on the following data representation
for billiard balls:

\begin{verbatim}
;; A BB is a:
;; (make-bb Color 
;;          (make-posn Integer Integer) 
;;          (make-vel Integer Integer))

;; A Color is one of:
;; - "red"
;; - "yellow"
;; - "green"

(define-struct bb (color center vel))
(define-struct vel (deltax deltay))
\end{verbatim}
The interpretation of a \verb|BB| is that the color is the color of
the ball, the \verb|posn| is the location of the center of the ball,
and the \verb|vel| structure describes the ball's velocity as a change
along the $x$-axis (\verb|deltax|) and $y$-axis (\verb|deltay|) in one
clock tick.

\noindent
Design a function called \verb|tock : BB -> BB| that calculates where
a given billiard ball will be, based on its velocity, after one tick
of the clock and assuming it does not encounter any obstacle.

\noindent
{\bf Problem 5 (10 points).} You've been hired by UPS, a package
delivery service, to manage and improve their truck routing software.
The software directs their truck drivers on how to proceed at each
intersection (we assume streets are laid out on a simple grid for this
problem): either continue straight, turn left, turn right, or take a
u-turn (i.e. turn around and proceed in the opposite direction.

You decide on the following data definition for representing these
instructions:

\begin{verbatim}
;; A Turn is one of:
;; - "S"
;; - "L"
;; - "R"
;; - "U"
;; Interp: directions for truck turns: straight (S), left (L), 
;; right (R), or u-turn (U).
\end{verbatim}

\noindent
{\bf Problem 5(a).} Write a template for \verb|Turn| functions.

@@

(define (turn-temp t)
  (cond [(string=? t "S") ...]
        [(string=? t "L") ...]
        [(string=? t "R") ...]
        [(string=? t "U") ...]))

@@

\noindent
{\bf Problem 5(b).} Write a stub for the following function:

\begin{verbatim}
;; backtrack : Turn -> Turn
;; Compute the turn to go backward on route
(check-expect (backtrack "S") "S")
(check-expect (backtrack "L") "R")
(check-expect (backtrack "R") "L")
(check-expect (backtrack "U") "U")
\end{verbatim}

@@

(define (backtrack t) "R")

@@

\noindent
{\bf Problem 6 (20 points).}
%
Part of your new UPS job is to compute useful metrics about trucking
routes, which will be used by the business office to optimize routes
for improving gasoline consumption.  A route consists of any number of
segments and each segment consists of a turn (as in problem 2) and a
distance to drive (measured in miles) after making that turn.

\vskip 1em
\noindent
Define a data definition for routes and give two examples of values
which are routes.

@@

\noindent
{\bf Problem 7 (20 points).}
%
Students seem obsessed with grades, which are calculated as numbers
between 0 and 100, inclusive.  At the end of the year, the professor
needs to figure out how to convert a numeric grade into a letter
grade, so they decide to write a small program using the following
data definition:
\begin{verbatim}
;; A Grade is one of:
;; - [90,100]  ;; Interp: A
;; - [80,90)   ;;         B
;; - [70,80)   ;;         C
;; - [60,70)   ;;         D
;; - [0,60)    ;;         F
\end{verbatim}

While students are obsessed with their individual grades, the
university is more interested in knowing statistics about the class as
a whole.  They ask the professor questions like ``how many `C'
students were in your class?,'' so they decide to write another
program using the following data definition:
\begin{verbatim}
;; A ClassGrades is one of:
;; - '()
;; - (cons Grade ClassGrades)
\end{verbatim}
But then, because the professor is lazy, they outsource the problem to
you.

\noindent
Design a program that calculates the number of `C' grades in a
\verb|ClassGrades| collection.

>>>>

\begin{center}
  {\Large {\bf CMSC 131A, Midterm 1}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Fall 2017 \ \\
  }
\end{center}

\vspace{2em}

\noindent
NAME:\verb|________________________________________________________|

\vspace{2em}

\noindent
UID: \verb|__________________________________________|

\vspace{2em}

\begin{center}
\begin{tabular}{| c | c |}
  \hline
  Question & Points \\ \hline \hline
  1 & 10 \\ \hline
  2 & 10 \\ \hline
  3 & 10 \\ \hline
  4 & 15 \\ \hline
  5 & 15 \\ \hline
  6 & 10 \\ \hline
  7 & 20 \\ \hline \hline
  Total: & 90\\
  \hline
\end{tabular}
\end{center}

@@

\noindent
Unless specifically instructed otherwise, you may use any built-in BSL
functions or data types.

@@

\noindent
{\bf Problem 1 (10 points).}
%
For the following program, write out each step of computation. At each
step, underline the expression being simplified. Label each step as
being ``arithmetic'' (meaning any built-in operation),
``conditional'', ``plug'' (for plugging in an argument for a function
parameter), or ``constant'' for replacing a constant with its value.

\begin{verbatim}
(define R 9)
(define (w z) (= R z))
(cond [(w 2) 4]
      [else 9])
\end{verbatim}

@@

\begin{SaveVerbatim}{VerbEnv}  
SOLUTION:
\end{SaveVerbatim}

%\censor{\vspace{-2.5in}}
\censor{%
\noindent\BUseVerbatim{VerbEnv}}

@@

\noindent
{\bf Problem 2 (10 points).}
%
For the following structure definition, list the names of every
function it creates.  For each function, classify it as being either a
constructor, accessor, or predicate.

\begin{verbatim}
(define-struct run (time up score))
\end{verbatim}

\noindent
{\bf Problem 3 (10 points).}
%
The graphical text editor you developed in a recent assignment has
attracted the interest of investors.  But before they write you a
check, they want to make sure you can handle fancy operations for
expert text editor users.  They ask if you can implement a \emph{swap}
operation, which will swap content to the left of the cursor with the
content to the right..  To satisfy your potential investors, you
define the following function for swapping strings at a given
position:
\begin{verbatim}
;; swap : String Integer -> String
;; Swap string to the left of i with string to the right.
;; Assume (<= 0 i (string-length s)).
(check-expect (swap "" 0) "")
(check-expect (swap "ab" 1) "ba")
(check-expect (swap "bannana" 3) "nanaban")
(define (swap s i) ...)
\end{verbatim}

\noindent
Give a correct definition for \verb|swap|.  (You only need to
provide code, not the design steps.)

(define (swap s i)
  (string-append (substring s i)
                 (substring s 0 i)))
\end{SaveVerbatim}

\newpage 
\noindent
{\bf Problem 4 (15 points).}
%
Use the following data definition for representing names:
\begin{verbatim}
;; A Name is a (make-name String String)
;; Interp: a person's full (first and last) name.
;; Both strings must contain at least one letter.
(define-struct name (first last))
\end{verbatim}

\noindent
Write a program, \verb|name-width|, that determines the width needed
to print someone's name on a letter using a mono-typed font (meaning
each letter has the same width).  The program can calculate this width
based on the length of the person's first and last name, plus the
width of a space to go in between.  Assume each letter is 10 pixels
wide.  For example, the name Bob Harper requires 100 pixels; the name
Bo Jack requires 70 pixels.  You may start from the following header:

\begin{verbatim}
;; name-width : Name -> Number
;; Computes width needed to print name in mono-type font
\end{verbatim}

@@

\newpage
\noindent
{\bf Problem 5 (15 points).}  
%
Design a program that takes a list of numbers and produces the count
of non-negative numbers in the list.  For example, if the list
contains 3, 4, -1, 0, -2, and 5, the count of non-negative numbers is
4.  You may use the following data definition:

\begin{verbatim}
;; A LoN (list of numbers) is one of:
;; - '()
;; - (cons Number LoN)
;; Interp: an arbitrarily long list of numbers
\end{verbatim}

@@

\noindent
{\bf Problem 6 (10 points).} You've been hired by the US Department of
Transportation to design traffic simulation software, which includes
simulating traffic lights.  Discovering the simple design from 131A is
insufficient, you come up with the following representation, which
makes it possible to model not just the color, but also how much
longer the current light has before changing:

\begin{verbatim}
;; A Light is one of:
;; - (make-red Natural)
;; - (make-yellow Natural)
;; - (make-green Natural)
;; Interp: color and remaining duration (in ticks) of a light 
(define-struct red (count))
(define-struct yellow (count))
(define-struct green (count))
\end{verbatim}
(Recall that a \verb|Natural| is a natural number, i.e. a non-negative
integer.)

\vspace{1em}

\noindent
{\bf Problem 6(a).} Write a template for \verb|Light| functions.

(define (light-temp l)
  (cond [(red? l) (... (red-count l) ...)]
        [(yellow? l) (... (yellow-count l) ...)]
        [(green? l) (... (green-count l) ...)]

@@

\noindent
{\bf Problem 6(b).} Write a stub for the following function:

\begin{verbatim}
;; tock : Light -> Light
;; Advance the light by one tick of time
(check-expect (tock (make-red 10)) (make-red 9))
(check-expect (tock (make-red 0)) (make-green 10))
\end{verbatim}

@@

(define (tock l) (make-red 0))

@@

\noindent
{\bf Problem 7 (20 points).}
%
Over the summer, you take an internship at the University of
Maryland's Center for Bioinformatics and Computational Biology (CBCB).
Your first task deals with computing information about \emph{DNA
  strands}.  A DNA strand is an arbitrarily long sequence of
\emph{nucleotides}, which are either: C, G, A, or T.  Design a program
that, given a DNA strand, computes the number of times the G
nucleotide occurs in it.  For example, if the strand is AGCTTTGA, your
program should produce 2.

;; SOLUTION:

;; A Nucleotide is one of:
;; - "C"
;; - "G"
;; - "A"
;; - "T"
;; Interp: DNA nucleotide C,G,A,T.

;; A DNA is one of:
;; - '()
;; - (cons Nucleotide DNA)
;; Interp: a DNA strand of nucleotides

;; count-g : DNA -> Natural
;; Count the number of occurrence of G in given DNA strand
(check-expect (count-g '()) 0)
(check-expect (count-g (cons "A" (cons "G" (cons "G" '())))) 2)
(define (count-g dna)
  (cond [(empty? dna) 0]
        [(cons? dna)
         (+ (if (string=? (first dna) "G") 1 0)
            (count-g (rest dna)))]))

@@

\noindent
[Extra space, should you need it.]

%% Design a program that consumes an arbitrarily large collection of
%% numbers and produces the count of non-negative numbers in the
%% collection.

%% \begin{SaveVerbatim}{VerbEnv}

%% SOLUTION:

%% ;; A LoN (list of numbers) is one of:
%% ;; - '()
%% ;; - (cons Number LoN)

%% ;; count-non-neg : LoN -> Natural
%% ;; Count the number of non-negative numbers in the given list.
%% (check-expect (count-non-neg '()) 0)
%% (check-expect (count-non-neg (cons -1 (cons 1 (cons -2 '())))) 1)
%% (define (count-non-neg lon)
%%   (cond [(empty? lon) 0]
%%         [(cons? lon) 
%%          (+ (if (<= 0 (first lon)) 1 0)
%%             (count-non-neg (rest lon)))]))
%% \end{SaveVerbatim}

%% \censor{%
%% \noindent
%% \BUseVerbatim{VerbEnv}}
